A Comparison of Parallel Graph Processing Implementations

### 2017-05-09 papers/publication/paper-pollard.tex
Updated methodology section in paper.

@@

Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation, profile Intel hardware counters, and measure the bottlenecks to measure scalability, computational efficiency for a variety of datasets.

@@

\subsection{Installing Libraries}
The libraries are stable forks of the given repositories which are configured to ensure the experiments execute in the same manner. Similarly, multiple copies of the datasets are stored to ensure the datasets are formatted correctly. For example, the roots for BFS are selected such that they are part of a nontrivial connected component.

@@

We refer to a graph's \emph{scale} when describing the size of the graph. Specifically, a graph with scale $S$ has $2^S$ vertices. For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex. The datasets are described further in Sect.~\ref{sec:datasets}. We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.

@@

The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League dataset\todo{One reviewer confused this dataset with the Kronecker.}. This dataset contains 61,670 vertices and 50,870,313 edges. This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}. This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.

@@

PageRank is selected because of its popularity; most libraries provide reference implementations. One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}. Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo{Better explain why these were selected. They're the most popular of the algorithms for the papers we surveyed.}

@@

\subsection{Parsing and Data Analysis}
The example workflow provided in our implementation yields the figures generated in this paper.

@@

In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately 33,554,432 edges)\todo{work?}. An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}. Graphalytics by default does not perform SSSP on unweighted, undirected graphs. The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo{What are the likely culprits for performance differences---measure cache miss rates. Try to find something which correlates with performance. Resource stalls will work easily (not as interesting) Cycles per instruction? Automate with perfexplorer or autoperf}

>>>>

Nai et al.~\cite{Nai:2016:architectural} provide a detailed performance analysis using GraphBIG as their reference implementation. Their approach also compares GraphBIG to the GraphLab and Pregel execution models. Their analysis considers architectural performance measurements such as cache miss rates to measure bottlenecks and computational efficiency for a variety of datasets. While incredibly thorough, these analyses focus on covering a wide range of datasets and appear difficult to apply to a new approach. For example, GraphMat reduce computation to sparse matrix operations which may not suffer from the same memory bottlenecks indicated in the results of Nai et al.

Beyond the aforementioned performance results there are a number of existing reports on performance analysis not tied to a particular software package we presently describe.

Satish et al.~\cite{Satish:2014:NavigatingGraph} analyze the performance of several systems on datasets on the order of $30$ billion edges and \cite{Lu:2014:ExperimentalEval} uses six real-world datasets and focuses on the vertex-centric programming model. These implementations are hand tuned and provides recommendations for future improvements.

@@

@@

\subsection{Installing Libraries}
The libraries are stable forks of the given repositories which are configured to ensure the experiments execute in the same manner.

@@

We refer to a graph's \emph{scale} when describing the size of the graph. Specifically, a graph with scale $S$ has $2^S$ vertices. For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex. We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our server, 72.

@@

\todo{More real world datasets}
The Graphalytics results in Table~\ref{tab:graphalytics} were performed on the Dota-League datasets. This dataset contains 61,670 vertices and 50,870,313 edges. This dataset comes from the Game Trace Archive\cite{Guo:2012:GTA} and is modified for Graphalytics\footnote{This dataset is available at \url{https://atlarge.ewi.tudelft.nl/graphalytics/}.}. This dataset is useful because it is both weighted and more dense than the usual real-world dataset with an average out-degree of 824.

@@

PageRank is selected because of its popularity; most libraries provide reference implementations. One challenge with using PageRank is the stopping criterion; all implementations have been modified to use $|| p_t - p_{t-1}||_1$ (the absolute sum of differences) where $p_t$ is the pagerank at step $t$. Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.\todo{Better explain why these were selected. They're the most popular of the algorithms for the papers we surveyed.}

@@

\subsection{Parsing and Data Analysis}
Our example analyzes data using the R programming language and shell scripting parsers. The example workflow provided in our utility yields the figures generated in this paper.

@@

In Table~\ref{tab:graphalytics} we show the results from running Graphalytics on a Kronecker graph of scale $22$ (4,194,304 vertices and approximately $16 \times 2^22 = 33,500,000$ edges)\todo{work?}. An explanation of each algorithm is given in \cite{Iosup:2016:Graphalyticstech}. Graphalytics by default does not perform SSSP on unweighted, undirected graphs. The discrepancy between PageRank values in Table~\ref{tab:graphalytics} and Fig.~\ref{fig:pr} is a result of the differing stopping criterion.\todo{What are the likely culprits for performance differences---measure cache miss rates. Try to find something which correlates with performance. Resource stalls will work easily (not as interesting) Cycles per instruction? Automate with perfexplorer or autoperf}
